---
title: 'Data Mining and Statistical Learning: Exercise 1'
author: "Milo Opdahl"
date: "February 10, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

```{r, include=FALSE}
library(tidyverse)
library(reshape2)
library(expss)

GB <- read.csv("D:/School Stuff/Graduate School/DMDS/ECO395M/data/greenbuildings.csv", header=TRUE)
view(GB)

# What were the results the "data guru" found?
#First, let's establish the subsets.
LeaseHigh <- GB[GB$leasing_rate > 10,]
Green <- GB[GB$green_rating==1,]
NotGreen <- GB[GB$green_rating !=1,]
median_rent <- median(LeaseHigh$Rent)
G_rent <- median(Green$Rent) # Median Rent for Green Buildings
NG_rent <- median(NotGreen$Rent) # Median Rent for Non-Green Buildings
D_rent <- G_rent-NG_rent #difference in rent from green and non-green
n <- 30 # Years of Operation
o_rate <- 0.90 # Occupancy Rate
sqft <- 250000 # Building Size
c_cost <- 100000000 # Building Cost
g_cert <- 0.05 # Green Certification

#Now, replicate the analysis.
payoff1 <- (c_cost*g_cert)/(D_rent*sqft) #at 100% occupancy
payoff2 <- (c_cost*g_cert)/(D_rent*sqft*o_rate) #at 90% occupancy

total1 <- (n-payoff1)*D_rent*sqft #total payoff under payoff1
total2 <- (n-payoff2)*D_rent*sqft*o_rate #total payoff under payoff2

full_rate <- c(100:11)/100 #Occupancy Rate Values >10
full_payoff <- (c_cost*g_cert)/(D_rent*sqft*full_rate) #Payoff Times based on Occupancy Rates

GB_plot <- data.frame(full_payoff,full_rate)

Guru_Analysis <- ggplot(GB_plot)+
  geom_point(mapping = aes(x = full_payoff, y = full_rate), color='black')+
  labs(title="Guru Payoff Analysis",subtitle="30 Year Operational Life",x="Years to Payoff",y="Occupancy Rate")+
  coord_cartesian(xlim=c(0, 30))+
  geom_vline(xintercept=7.7, linetype="solid", color = "red")+
  geom_hline(yintercept=0.25, linetype="dashed", color = "red")+
  geom_vline(xintercept=30, linetype="solid", color = "red")

```

## "Data Guru" Analysis


```{r Guru_Analysis, echo=FALSE}
Guru_Analysis
```

  The so-called "Data Guru" likely came to their conclusions similar to the chart above, utilizing certain methods such as a median rent measure of green vs non-green buildings and removing data on low-occupancy buildings under the theory that such information could "potentially distort the analysis."  From this we can see how they came to their conclusions on having a total payoff and breaking even on the investment after about 7.7 years, assuming 100% occupancy rate.  However, their approach has major faults, particularly in two matters.
  
* By removing data on low-occupancy buildings, we lose potentially important insight on market conditions for both green and non-green buildings, such as their rental values and knowing what type of buildings are more likely to have lower occupancy rates.

* Utilizing a median rent measure instead of a mean rental valuation of green vs non-green buildings may allow to control for robustness in outliers; however, controlling for robustness in outliers will skew the payoff data that is being gathered, specifically by overestimating the effect of green buildings.  

```{r echo=FALSE}
table <- matrix(c(
  median(LeaseHigh$Rent),
  mean(LeaseHigh$Rent),
  median(Green$Rent),
  mean(NotGreen$Rent),
  median(Green$Rent),
  mean(NotGreen$Rent)
),ncol=2,byrow=TRUE)

colnames(table) <- c("Median","Mean")
rownames(table) <- c("Whole Dataset","Green","Non-Green")

table <- as.table(table)
table
```

  The table above shows that there is a 2.6 difference using median-values, while there will only be a 1.7 difference in mean-values.  By using mean-values, the break-even point will be extended further than the "Data Guru" would anticipate.
  
## A New Analysis

  By using the full data and mean rent values, a new (and arguably better) understanding of the data can be found.
```{r New_Analysis, echo=FALSE}
GB <- read.csv("D:/School Stuff/Graduate School/DMDS/ECO395M/data/greenbuildings.csv", header=TRUE)

#Subsets
Green <- GB[GB$green_rating==1,]
NotGreen <- GB[GB$green_rating !=1,]
mean_rent <- mean(GB$Rent) 
G_rent <- mean(Green$Rent) # Mean Rent for Green Buildings
NG_rent <- mean(NotGreen$Rent) # Mean Rent for Non-Green Buildings
D_rent <- G_rent-NG_rent #difference in rent from green and non-green
n <- 30 # Years of Operation
o_rate <- 0.90 # Occupancy Rate
sqft <- 250000 # Building Size
c_cost <- 100000000 # Building Cost
g_cert <- 0.05 # Green Certification

#Analysis
payoff1 <- (c_cost*g_cert)/(D_rent*sqft) #at 100% occupancy
payoff2 <- (c_cost*g_cert)/(D_rent*sqft*o_rate) #at 90% occupancy

total1 <- (n-payoff1)*D_rent*sqft #total payoff under payoff1
total2 <- (n-payoff2)*D_rent*sqft*o_rate #total payoff under payoff2

full_rate <- c(100:1)/100 #Occupancy Rate Values
full_payoff <- (c_cost*g_cert)/(D_rent*sqft*full_rate) #Payoff Times based on Occupancy Rates

GB_plot <- data.frame(full_payoff,full_rate)

New_Analysis <-ggplot(GB_plot)+
  geom_point(mapping = aes(x = full_payoff, y = full_rate), color='black')+
  labs(title="New Payoff Analysis",subtitle="30 Year Operational Life",x="Years to Payoff",y="Occupancy Rate")+
  coord_cartesian(xlim=c(0, 30))+
  geom_vline(xintercept=payoff1, linetype="solid", color = "red")+
  geom_hline(yintercept=0.37, linetype="dashed", color = "red")+
  geom_vline(xintercept=30, linetype="solid", color = "red")
New_Analysis
```

  Here is a more accurate estimate of the payoff timetable for a green building using mean-values. For 100% occupancy, the payoff would take place within 11.43 years. In addition, at the building's estimated minimal lifecycle the payoff would take place given a 37% occupacy rate, with anything lower resulting in a loss.
  
\newpage

## Question 2

```{r ABIA, include=FALSE}
library(plyr)
library(ggplot2)
library(tidyverse)
library(reshape2)
library(expss)

ABIA = read.csv("D:/School Stuff/Graduate School/DMDS/ECO395M/data/ABIA.csv", header=TRUE)
View(ABIA)

flight1 <- ABIA[1:4]
flight2 <- ABIA[9]
flight3 <- ABIA[19]
ABIAf <- cbind(flight1, flight2, flight3)
head(ABIAf)

MilesperMin <- ABIA$Distance / ABIA$ActualElapsedTime
ABIAf <- cbind(ABIAf, MilesperMin)
head(ABIAf)

ABIAf <- subset(ABIAf, UniqueCarrier == "AA")
ABIAf <- subset(ABIAf, !is.na(Distance))
ABIAf <- subset(ABIAf, !is.na(MilesperMin))
head(ABIAf)

summary(ABIAf)

ggplot(ABIAf, aes(y=MilesperMin, x=DayOfWeek))+geom_violin()

plot1 <- ggplot(ABIAf, aes(x = DayOfWeek, y = MilesperMin)) + geom_point(aes(color = ABIAf$DayofMonth)) + facet_wrap(~Month)

```  
# American Airline Flights At ABIA

The following dot plot gives the speed of American Airline flights to and from ABIA per day of the week, in every month of 2008.  By getting to understand this graph, we can figure out what days have the fastest flight times or have the highest density of fast flights.

```{r plot_1,echo=FALSE}
plot1
```

\newpage
  
## Question 3

```{r load data,include=FALSE}
library(tidyverse)
library(FNN)

sclass = read.csv("D:/School Stuff/Graduate School/DMDS/ECO395M/data/sclass.csv", header=TRUE)
sclass <- select(sclass,trim,mileage,price)

# The variables involved

# plot the data
#ggplot(data = sclass) + 
#  geom_point(mapping = aes(x = mileage, y = price), color='darkgrey')
#summary(sclass)

# Focus on 2 trim levels: 350 and 65 AMG
sclass350 = subset(sclass, trim == '350')

sclass65AMG = subset(sclass, trim == '65 AMG')

# create a function to automate everything

function_knn <- function(x){
  # Make a train-test split
  N = nrow(x)
  N_train = floor(0.8*N)
  N_test = N - N_train
  
  #####
  # Train/test split
  #####
  
  # randomly sample a set of data points to include in the training set
  train_ind = sample.int(N, N_train, replace=FALSE)
  
  # Define the training and testing set
  D_train = x[train_ind,]
  D_test = x[-train_ind,]
  
  # optional book-keeping step:
  # reorder the rows of the testing set by the mileage
  D_test = arrange(D_test, mileage)
  head(D_test)
  
  # Now separate the training and testing sets into features (X) and outcome (y)
  X_train = select(D_train, mileage)
  y_train = select(D_train, price)
  X_test = select(D_test, mileage)
  y_test = select(D_test, price)
  
  
  #####
  # Fit a few models
  #####
  
  # linear and quadratic models
  lm1 = lm(price ~ mileage, data=D_train)
  lm2 = lm(price ~ poly(mileage, 2), data=D_train)
  
  # KNN
  knn = knn.reg(train = X_train, test = X_test, y = y_train, k=knn_size)
  
  #####
  # Compare the models by RMSE_out
  #####
  
  # define a helper function for calculating RMSE
  rmse = function(y, ypred) {
    sqrt(mean(data.matrix((y-ypred)^2)))
  }
  
  ypred_lm1 = predict(lm1, X_test)
  ypred_lm2 = predict(lm2, X_test)
  ypred_knn = knn$pred
  
  rmse(y_test, ypred_lm1)
  rmse(y_test, ypred_lm2)
  rmse(y_test, ypred_knn)
  
  
  ####
  # plot the fit
  ####
  
  # attach the predictions to the test data frame
  D_test$ypred_lm2 = ypred_lm2
  D_test$ypred_knn = ypred_knn
  
  p_test = ggplot(data = D_test) + 
    geom_point(mapping = aes(x = mileage, y = price), color='lightgrey') + 
    theme_bw(base_size=18)
  p_test
  
  p_test + geom_path(aes(x = mileage, y = ypred_knn), color='red')+ 
    geom_path(aes(x = mileage, y = ypred_lm2), color='blue')+
    ggtitle(paste("S-Class",trim,"when KNN=",knn_size))
}
```

# S-Class 350

```{r plot 1,echo=FALSE}
trim <- 350

knn_size <- 10
function_knn(sclass350)

knn_size <- 20
function_knn(sclass350)

knn_size <- 40
function_knn(sclass350)

knn_size <- 50
function_knn(sclass350)
```

\newpage

# S-Class 65 AMG

```{r plot 2,echo=FALSE}
trim <- "65 AMG"

knn_size <- 10
function_knn(sclass65AMG)

knn_size <- 20
function_knn(sclass65AMG)

knn_size <- 40
function_knn(sclass65AMG)

knn_size <- 50
function_knn(sclass65AMG)
```
